Comment on Hodgkins et al. (2018)

### 2023-11-27 analysis/paper/002-reply-main.Rmd
Update the comment on Hodgkins et al. (2018) with comments from Klaus.

1. Larger aromatic contents may be necessary to stabilize tropical near-surface peat against decomposition and aromatics may accumulate at a higher rate with depth in tropical peatlands than estimated by @Hodgkis.2018.

@@

Our previous study has shown that in particular the spectral prediction model for Klason lignin is biased, especially for more decomposed peat [@Teickner.2022f]. Our reanalysis using the modified model for Klason lignin from @Teickner.2022f thus indicates that larger Klason lignin contents may be necessary to stabilize tropical near-surface peat against decomposition as well as that changes in Klason lignin with depth may be more pronounced in the tropics than estimated in @Hodgkins.2018 (supporting information S1).

@@

We argue that we should differentiate between vague concepts such as carbohydrates and aromatics and precise concepts such as holocellulose and Klason lignin and we should always use the most precise concept possible, unless more general statements are explicitly warranted.

For example, limitations of the extraction procedure by which holocellulose and Klason lignin are defined are known (e.g., @Hatfield.2005, @Bunzel.2011, @Abu-Omar.2021) and these limitations as well as differences to other procedures get obscured by using vague words such as carbohydrates and aromatics.

Using precise concepts for organic matter fractions becomes especially important when interpreting peat chemistry based on mid infrared spectra (MIRS) or spectral prediction models because all variables derived from MIRS --- for example the widely used humification indices, as proxy for relative contents of recalcitrant organic matter fractions (e.g. @Broder.2012), or Klason lignin contents predicted from models --- are really only indirect estimates of organic matter fractions and there are abundant possibilities how they can be misleading if they are not sufficiently validated.

Intensities recorded in MIRS really only represent the fraction of the incident infrared radiation which is absorbed by specific molecular structures which happen to absorb at that specific energy level [@Stuart.2004]. Such molecular structures may for example be aromatic C=C bonds, C-N bonds in amides, or C-O bonds in alcohols (e.g. carbohydrates) [@Stuart.2004]. In addition, absorption of infrared radiation (and hence peak heights) can be decoupled from abundances of molecular structures because also the electronegativity of the bonds controls absorption of mid infrared radiation [@Stuart.2004].

One major reason why the model for Klason lignin in @Hodgkins.2018 is biased and invalid for peat is that the only predictor variable used in this model, `arom15arom16`, is not a good indicator for Klason lignin because also proteins can absorb in the same energy range [@Stuart.2004] and therefore, estimates are biased relative to the amount of proteins in the samples [@Teickner.2022f]. Therefore, we should interpret `arom15arom16` and similar variables, such as specific humification indices (HI$_{1630/1090}$, HI$_{1510/1090}$) only then as good proxies for aromatics if we have shown that protein contents do not differ much between the peat samples (thresholds which still needs to be established).

In all other cases, we should be precise in our wording by calling the variable `arom15arom16` to signal that this variable may be no good indicator for aromatics if it is confounded by proteins. Obviously, this recommendation equally applies to all other variables derived from MIRS.

>>>>

1. Larger aromatic contents may be necessary to stabilize tropical near-surface peat against decomposition and aromatics may accumulate at a higher rate with depth in tropical peatlands than estimated by @Hodgkins.2018.

@@

Our previous study has shown that the spectral prediction model for Klason lignin is biased, especially for more decomposed peat [@Teickner.2022f]. Our reanalysis using the modified model for Klason lignin from @Teickner.2022f indicates that larger Klason lignin contents may be necessary to stabilize tropical near-surface peat against decomposition as well as that changes in Klason lignin with depth may be more pronounced in the tropics than estimated in @Hodgkins.2018 (supporting information S1).

@@

We argue that we should differentiate between vague concepts such as carbohydrates and aromatics and precise concepts such as holocellulose and Klason lignin and we should always use the most precise concept possible to describe the organic matter fraction we _plan_ (or intent) to measure, unless more general statements are explicitly warranted.

For example, @Hodgkins.2018 planned to measure Klason lignin contents because their spectral prediction model used Klason lignin data as dependent variable, however this variable is labeled as aromatics. Limitations of the extraction procedure by which Klason lignin are defined are known (e.g., @Hatfield.2005, @Bunzel.2011, @Abu-Omar.2021) and these limitations as well as differences to other procedures get obscured by using vague words such as carbohydrates and aromatics.

However, using precise concepts for organic matter fractions is also important for exactly the opposite reason: to not suggest a variable has more precision than is actually the case; this becomes especially important when interpreting peat chemistry based on mid infrared spectra (MIRS) or spectral prediction models because all variables derived from MIRS --- for example the widely used humification indices, as proxy for relative contents of recalcitrant organic matter fractions (e.g. @Broder.2012), or Klason lignin contents predicted from models --- are in fact only indirect estimates of organic matter fractions and there are abundant possibilities how they can be misleading if they are not sufficiently validated. Stating these limitations requires precise concepts.

Intensities recorded in MIRS really only represent the fraction of the incident infrared radiation which is absorbed by specific molecular structures which happen to absorb at that specific energy level [@Stuart.2004]. Such molecular structures may for example be aromatic C=C bonds, C-N bonds in amides, or C-O bonds in alcohols (e.g. carbohydrates) [@Stuart.2004]. In addition, absorption of infrared radiation (and hence peak heights or areas) can be decoupled from abundances of molecular structures because also the electronegativity of the bonds controls absorption of mid infrared radiation [@Stuart.2004]. These confounding factors make it necessary to explicitly define the conditions under which a spectral prediction model or other variable derived from MIRS (e.g. a humification index) is a valid proxy for a specific, precisely defined, organic matter fraction.

One major reason why the model for Klason lignin in @Hodgkins.2018 is biased and invalid for peat is that the only predictor variable used in this model, `arom15arom16`, is not a good indicator for Klason lignin because also proteins can absorb in the same energy range [@Stuart.2004] and therefore, estimates are biased depending on the amount of proteins in the samples [@Teickner.2022f]. Therefore, we should interpret `arom15arom16` and similar variables, such as specific humification indices (HI$_{1630/1090}$, HI$_{1510/1090}$) only then as good proxies for the relative abundance of aromatic C=C bonds if we have shown that protein contents do not differ much between the peat samples (thresholds which still need to be established).

In all other cases, where we cannot validate if a variable is a good proxy for a specific organic matter fraction, we should be precise in our wording by calling the variable `arom15arom16` (instead of Klason lignin or aromatics) to signal that this variable may be no good indicator for aromatics if it is confounded by proteins, i.e. we may agree on `arom15arom16` as precise name for this variable, but it must always be understood that it is first and foremost _only_ defined as sum of the area of two peaks extracted by the specific procedure proposed in @Hodgkins.2018 and nothing more. Obviously, this recommendation equally applies to all other variables derived from MIRS.
